---
title: "Codebook for Project"
author: "Robert Richardson"
date: "December 14, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Description of script


This script uses data from an experiment that tracked activity recognition using smartphones. The experiments have been carried out with a group of 30 volunteers within an age bracket of 19-48 years. Each person performed six activities (WALKING, WALKING_UPSTAIRS, WALKING_DOWNSTAIRS, SITTING, STANDING, LAYING) wearing a smartphone (Samsung Galaxy S II) on the waist. The script begins by reading in the data.

```{r}
library(dplyr)
features <- read.table("data/features.txt", col.names = c("n","functions"))
activities <- read.table("data/activity_labels.txt", col.names = c("code", "activity"))
subject_test <- read.table("data/test/subject_test.txt", col.names = "subject")
x_test <- read.table("data/test/X_test.txt", col.names = features$functions)
y_test <- read.table("data/test/y_test.txt", col.names = "code")
subject_train <- read.table("data/train/subject_train.txt", col.names = "subject")
x_train <- read.table("data/train/X_train.txt", col.names = features$functions)
y_train <- read.table("data/train/y_train.txt", col.names = "code")
```


For the purposes of this project, a subset of the original dataset was obtained by merging two existing files. One contained data from a training run and the other captured data from the actual testing. 

```{r}

X <- rbind(x_train, x_test)
Y <- rbind(y_train, y_test)
Subject <- rbind(subject_train, subject_test)
Merged_Data <- cbind(Subject, Y, X)
```


After merging the data sets, the script extracts mean and standard deviation measures for each measurement and uses descriptive activity names to name the activities in the new dataset.

```{r}
SD_M_Data <- Merged_Data %>% select(subject, code, contains("mean"), contains("std"))
SD_M_Data$code <- activities[SD_M_Data$code, 2]

names(SD_M_Data)[2] = "activity"
names(SD_M_Data)<-gsub("Acc", "Accelerometer", names(SD_M_Data))
names(SD_M_Data)<-gsub("Gyro", "Gyroscope", names(SD_M_Data))
names(SD_M_Data)<-gsub("BodyBody", "Body", names(SD_M_Data))
names(SD_M_Data)<-gsub("Mag", "Magnitude", names(SD_M_Data))
names(SD_M_Data)<-gsub("^t", "Time", names(SD_M_Data))
names(SD_M_Data)<-gsub("^f", "Frequency", names(SD_M_Data))
names(SD_M_Data)<-gsub("tBody", "TimeBody", names(SD_M_Data))
names(SD_M_Data)<-gsub("-mean()", "Mean", names(SD_M_Data), ignore.case = TRUE)
names(SD_M_Data)<-gsub("-std()", "STD", names(SD_M_Data), ignore.case = TRUE)
names(SD_M_Data)<-gsub("-freq()", "Frequency", names(SD_M_Data), ignore.case = TRUE)
names(SD_M_Data)<-gsub("angle", "Angle", names(SD_M_Data))
names(SD_M_Data)<-gsub("gravity", "Gravity", names(SD_M_Data))
```


Finally, a second, independent tidy dataset with the average of each variable for each activity and each subject is created and written to disk.

```{r}
FinalData <- SD_M_Data %>%
  group_by(subject, activity) %>%
  summarise_all(funs(mean))
write.table(FinalData, "FinalData.txt", row.name=FALSE)
```


